\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2017-10-25 document/XSGRB_sample.tex
further work on intro and conclusions

The paper is
organized in the following way: In Sect.~\ref{sample} we describe the sample
including the sample selection criteria and the observational strategy.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
towards about 10$^5$ QSOs have a handful such systems with
$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
\citep[e.g., five in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, 
reveal such
systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
Whereas DLAs
towards QSOs are mostly limited to $1.8\lesssim z \lesssim 5$ due to the
atmospheric  UV-cutoff and increasing Lyman-blanketing at increasing redshifts
\citep[e.g.,][]{Rafelski2014}, GRBs allow us to see into the hearts of
star-forming galaxies over the full history of cosmic star formation from $z
\approx 0$ to $z > 8$ \citep[e.g.,][]{Tanvir2009, Salvaterra2009,
    Jakobsson2012}. With afterglow spectroscopy (throughout the electromagnetic
spectrum from X-rays to the sub-mm) we can hence characterize the properties of
star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
molecular contents, ISM temperatures, UV-flux densities, etc.

Bias still, more to do, ELTs, SVOM unharvested potential Kruehler? 2012, GRB08607

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{acknowledgements}
We should add an ode to Neil Gehrels here.
JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
acknowledges support by the European Commission under the Marie Curie
Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
European Commission under the Marie Curie Career Integration Grant programme
(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
acknowledge expert support from the ESO staff at the Paranal and La Silla
observatories in obtaining these target of opportunity data.

>>>>

The paper is organized in the following way: In Sect.~\ref{sample} we describe
the sample including the sample selection criteria and the observational
strategy.  In Sec.t~\ref{results} we describe the results of the survey, i.e.
the efficiency of the follow-up effort and the characteristics of the observed
bursts. We also here assess the completeness of the finally realized sample.
Finally, we offer our conclusions in Sect.~\ref{conclusions}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and conclusions}\label{conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Our sample serves the purpose to further characterize the environments 
of GRBs that was also much advanced by the previous surveys based primarily on 
lower-resolution spectographs. The most striking result found is that GRBs
typically are located behind very large column densities of neutral hydrogen.
Only after observing more than 12000 damped Lyman-$\alpha$ absorbers (DLAs)
towards about 10$^5$ quasars have a handful such systems with
$\log({N_\mathrm{HI}/\mathrm{cm^{-2}}}) > 22$ been identified
\citep[e.g., five in][]{Noterdaeme2012}. Long GRB afterglow spectra, by contrast, 
reveal such
systems in the majority of cases \citep[][and this work]{Jakobsson2006, Fynbo2009}. 
With afterglow spectroscopy (throughout the electromagnetic
spectrum from X-rays to the sub-mm) we hence characterize the properties of
star-forming galaxies over cosmic history in terms of  redshifts, metallicities,
molecular contents, ISM temperatures, UV-flux densities, etc.

It seems important to end this paper by stressing that the potential of 
using GRB sightlines as probes is far from fully harvested. The sample of
sightlines probed by our spectra is not representative for all GRB 
sightlines as we have shown (CONFIRM) and consistent with earlier findings
\citep[e.g.,][]{Fynbo2009}. \cite{Kruhler2013} convincingly argue, that 
very rich sightlines like that probed by the remarkable GRB\,080607 
\citep{Prochaska09} are probably significantly more frequent than in the sightlines 
sampled by our spectra. However, with current instrumentation these sightlines are
out of reach except under very fortunate circumstances as in the case of 
GRB\,080607 when the afterglow could be observed only a few minutes after the 
burst with a 10-m class telescope. Observations of such sightlines with 
X-shooter-like spectrographs on th enext generation of 20--40-m telescopes
is likely to be very rewarding.

\begin{acknowledgements}
*** We should add an ode to Neil Gehrels here ***.
JPUF, BMJ and DX acknowledge support from the ERC-StG grant EGGS-278202.  The
Dark Cosmology Centre is funded by the Danish National Research Foundation.  TK
acknowledges support by the European Commission under the Marie Curie
Intra-European Fellowship Programme in FP7.  AdUP acknowledges support by the
European Commission under the Marie Curie Career Integration Grant programme
(FP7-PEOPLE-2012-CIG 322307).  This work made use of data supplied by the UK
{\it Swift} Science Data Centre at the University of Leicester.  Finally, we
acknowledge expert support from the ESO staff at the Paranal and La Silla
observatories in obtaining these target of opportunity data.
